---
title: "ML_Assignment_2"
author: "SrushtiP"
date: "10/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries required for the assignment.

```{r}
library(ISLR)
library(caret)
library(dummies)
library(FNN)
library(gmodels)
options(warn=-1)
```

To read the data

```{r}
UniversalBank <- read.csv("UniversalBank.csv")
```

Create dummy for the Education column

```{r}
Education_Dummy <- dummy(UniversalBank$Education)

temp<- cbind(UniversalBank, Education_Dummy)

UniversalBank_Object <- temp[c(-1,-5, -8)]
head(UniversalBank_Object)
```

Set the seed for partiotion in an orderly manner

```{r}
set.seed(15)
```

Partition Data in 60/40 % format

```{r}
Training_Data_Index <- createDataPartition(UniversalBank_Object$Age, p = 0.6, list = FALSE)

Training_Data <- UniversalBank_Object[Training_Data_Index,]
dim(Training_Data)

Validation_Data <- UniversalBank_Object[-Training_Data_Index,]
dim(Validation_Data)
```

Normalization

```{r}
norm.values <- preProcess(Training_Data, method = c ("center", "scale"))

train.norm.df <- predict(norm.values, Training_Data)

Valid.norm.df <- predict(norm.values, Validation_Data)

summary(train.norm.df)
summary(Valid.norm.df)
```

Differentiate Predictors and Prediction data 

```{r}
Train_Data <- train.norm.df[,-7]
Train_Outcome <- factor(Training_Data[,7], levels = c(0,1), labels = c("Deny","Accept"))

Valid_Data <- Valid.norm.df[,-7]
Valid_Outcome <- factor(Validation_Data[,7], levels = c(0,1), labels = c("Deny", "Accept"))
```

Apply KNN with k value as 1 for Validation and Test data

```{r}
knn_obj <- knn(Train_Data, Valid_Data, cl=Train_Outcome, k=1, prob = TRUE)
View(knn_obj)
```

Q. 1.)
Test KNN on test data

```{r}
Test_Data <- c(40, 10, 84, 2, 2, 0, 0, 0, 1, 1, 0, 1, 0)

knn_test <- knn(Train_Data, Test_Data, cl=Train_Outcome, k=1,prob = TRUE)
knn_test
```

Q. 2.)
Hypertuning

```{r}
accuracy_data <- data.frame(k = seq(1,55,1), accuracy = rep (0, 55))

for(i in 1:55){
  knn_pred <- knn(Train_Data, Valid_Data, cl=Train_Outcome, k=i)
  accuracy_data[i,2] <- confusionMatrix(knn_pred, Valid_Outcome)$overall[1]
}
accuracy_data

best_k <- accuracy_data[which.max(accuracy_data$accuracy),]
best_k
```

Q. 3.)
Create Confusion Matrix

```{r}
knn_test_k <- knn(Train_Data, Valid_Data, cl=Train_Outcome, k=best_k$k, prob = TRUE)
View(knn_test_k)

CrossTable(x = Valid_Outcome, y = knn_test_k, prop.chisq = FALSE)
```

Accuracy is 96.64 %

Q. 4.)
Test the knn model using hypertuned k value for tet data.

```{r}
BEST_KNN_TEST <- knn(Train_Data, Test_Data, cl=Train_Outcome, k=, prob = TRUE)
BEST_KNN_TEST
```

Q. 5.)
Partition Data i 50-30-20 % format

```{r}
Test_Data_Index_2 <- createDataPartition(UniversalBank_Object$Age, p = 0.2, list = FALSE)

Testing_Data_2 <- UniversalBank_Object[Test_Data_Index_2,]

Travel_Data_2 <- UniversalBank_Object[-Test_Data_Index_2,]

Training_Data_Index_2 <- createDataPartition(Travel_Data_2$Age, p = 0.625, list = FALSE)

Training_Data_2 <- Travel_Data_2[Training_Data_Index_2,]

Validation_Data_2 <- Travel_Data_2[-Training_Data_Index_2,]

dim(Training_Data_2)
dim(Validation_Data_2)
dim(Testing_Data_2)
```

Normalization

```{r}
norm.values <- preProcess(Training_Data_2, method = c ("center", "scale"))

train.norm.df_2 <- predict(norm.values, Training_Data_2)

Valid.norm.df_2 <- predict(norm.values, Validation_Data_2)

test.norm.df_2 <- predict(norm.values, Testing_Data_2)
```

Differentiate Predictors and Prediction data 

```{r}
Train_Data_2 <- train.norm.df_2[,-7]
Train_Outcome_2 <- factor(Training_Data_2[,7], levels = c(0,1), labels = c("Deny","Accept"))

Valid_Data_2 <- Valid.norm.df_2[,-7]
Valid_Outcome_2 <- factor(Validation_Data_2[,7], levels = c(0,1), labels = c("Deny", "Accept"))

Test_Data_2 <- test.norm.df_2[,-7]
Test_Outcome_2<- factor(Testing_Data_2[,7], levels = c(0,1), labels = c("Deny", "Accept"))
```

Apply KNN on Training and Validation set

```{r}
knn_Validation <- knn(Train_Data_2, Valid_Data_2, cl=Train_Outcome_2, k=best_k$k, prob = TRUE)
View(knn_obj)

CrossTable(x = Valid_Outcome_2, y = knn_Validation, prop.chisq = FALSE)

knn_Testing <- knn(Train_Data_2, Test_Data_2, cl=Train_Outcome_2, k=best_k$k,prob = TRUE)
knn_test

CrossTable(x = Test_Outcome_2, y = knn_Testing, prop.chisq = FALSE)
```

The Accuracy for the Validation Data is: 95.79439252336449 %

The Accuracy for the Testing Data is: 95.004995004995 %