---
title: "Machine Learning Assignment 1"
author: "SrushtiP"
date: "10/27/2019"
output: html_document
---

Problem Statement:

This assignment will concentrate on using R and Git. Specifically, you will do the following:

- Download a dataset from the web. You may use any source, but specify the source in your code. 
  Also ensure that the data has a mix of quantitative and qualitative (categorical) variables.
- Import the dataset into R
- Print out descriptive statistics for a selection of quantitative and categorical variables.
- Transform at least one variable. It doesn't matter what the transformation is.
- Plot at least one quantitative variable, and one scatterplot

* Source of DataSet: 

* https://dasl.datadescription.com/datafiles/?_sfm_cases=4+59943&sort_order=modified+desc&sf_paged=9


To import the dataset in R.

```{r}
Stream_Statistics <- read.csv("Streams.csv")
```

To inspect your data: head() Prints the 1st 5 entries of each column and tail() Prints the last 5 entries of each column.

```{r}
head(Stream_Statistics, 5)

tail(Stream_Statistics, 5)
```

* Descriptive Statistics of the dataset

Summmary() helps describe the data statistics of each column.

If the column is a numeric variable, mean, median, min, max and quartiles are returned.
If the column is a factor variable, the number of observations in each group is returned.

```{r}
summary(Stream_Statistics)
```

1. Mean : The Arithmetic Mean is the average of the numbers or a calculated "central" value of a set of numbers.

```{r}
mean(Stream_Statistics$pH)
```

2. Median : The middle value. It’s a robust alternative to mean.

```{r}
median(Stream_Statistics$Temp)
```

3. Mode : The most frequent value

```{r}
require(modeest)
mfv(Stream_Statistics$pH)
```

4. Variance : The variance represents the average squared deviation from the mean.

```{r}
var(Stream_Statistics$Hard)
```

5. Standard Deviation : The standard deviation is the square root of the variance. It measures the average deviation of the values, in the data, from the mean value.

```{r}
sd(Stream_Statistics$Alkali)
```

6. Absolute Deviation : The median absolute deviation (MAD) measures the deviation of the values, in the data, from the median value.

```{r}
mad(Stream_Statistics$Phosphate)
```

7. Minimum Value : Provides the minimum value of the specified column

```{r}
min(Stream_Statistics$Hard)
```

8. Maximum Value : Provides the maximum value of the specified column

```{r}
max(Stream_Statistics$Alkali)
```

9. Which is the row with minimum of the desired column

```{r}
Stream_Statistics[which.min(Stream_Statistics$Hard),]
```

10. Which is the row with maximum of the desired column

```{r}
Stream_Statistics[which.max(Stream_Statistics$Alkali),]
```

11. Range: Range corresponds to biggest value minus the smallest value. It gives you the full spread of the data.

```{r}
range(Stream_Statistics$Temp)
```

12. Quantile : quartiles divide the data into 4 parts.

The interquartile range (IQR) - corresponding to the difference between the first and third quartiles - is sometimes used as a robust alternative to the standard deviation.

```{r}
quantile(Stream_Statistics$Phosphate)
IQR(Stream_Statistics$Phosphate)
```

* Transformation

Transforming data is one step in addressing data that do not fit model assumptions, and is also used to coerce different variables to have similar distributions.

rcompanion- Library require for function plotNormalHistogram

Histogram for pH value

```{r}
library(rcompanion)

plotNormalHistogram(Stream_Statistics$pH, main = "Normal pH")
```

1. Square Root Transformation

```{r}
Sq_Trans = sqrt(Stream_Statistics$pH)
print("Square transformation of pH")
print(head(Sq_Trans), 5)

plotNormalHistogram(Sq_Trans, main = "Square Root Transformed pH")
```

2. Cube Root Transformation

```{r}
C_Trans = abs(Stream_Statistics$pH)^(1/3)
print("Cube Transformation of pH")
print (head(C_Trans),5)

plotNormalHistogram(C_Trans, main = "Cube Root Transformed pH")
```

3. Log Transformation

```{r}
log_Trans = log(Stream_Statistics$pH)
print("Log Transformation of pH")
print(head(log_Trans), 5)

plotNormalHistogram(log_Trans, main = "Log Root Transformed pH")
```

* Graphical Representation

1. Boxplot

```{r}
plot(Stream_Statistics$Substrate, Stream_Statistics$pH, xlab = "Substrate", ylab = "pH", main = "Substrate vs pH of Streams")
```

2. Bar Chart - Plotted a graph of number of different substrate

```{r}
barplot(table(Stream_Statistics$Substrate), xlab = "Substrate", ylab = "Count", main = "Count of Substrate")
```

3. Scatter Plot

```{r}
plot.default(Stream_Statistics$pH, Stream_Statistics$Hard, xlab = "pH of Stream", ylab = "Hardness", main = "pH vs Hardness of different Stream Water")
```